### 2024-03-06 paper/main.tex
Add comment about warm-up phase.
Add description of baseline

We benchmark the effectiveness of such a parallelization scheme by implementing Alg.~\ref{alg:fast} in \texttt{C++} using vectorization via SIMD instructions (CPU) and a CUDA implementation (GPGPU) and compare the performance with implementations of Alg.~\ref{alg:no_recursion} and \ref{alg:linear}.
Note that we intentionally not pick Alg.~\ref{alg:vanilla} for reference as this variant crashes due to its recursion nature quickly for larger trajectory sizes.

@@

All implementations use 32-bit floating point numbers to represent (intermediate) steps during calculation and run on an Ubuntu 22.04 machine using GCC-10 for the CUDA variant and GCC-11 for the rest.
The SIMD variant uses the \texttt{C++} extensions for parallelism as implemented in GCC-11~\citep{N4808}.
Builds where optimized using the flags \texttt{-O3}, \texttt{-DNDEBUG}, \texttt{-march=native}, and \texttt{-ffast-math}.

@@

\begin{figure}[htbp]
    \centering
    \begin{subfigure}{.49\textwidth}
        \includegraphics[width=.8\textwidth]{img/laptop/abs_performance-N.png}
        \caption{Absolute runtime with $P = 2^{10}$.}
    \end{subfigure}
    \begin{subfigure}{.49\textwidth}
        \includegraphics[width=.8\textwidth]{img/laptop/abs_performance-pP.png}
        \caption{Absolute runtime with $N = 2^{10}$.}
    \end{subfigure}\\
    \begin{subfigure}{.49\textwidth}
        \includegraphics[width=.8\textwidth]{img/laptop/rel_performance-N.png}
        \caption{Relative improvement with $P = 2^{10}$.}
    \end{subfigure}
    \begin{subfigure}{.49\textwidth}
        \includegraphics[width=.8\textwidth]{img/laptop/rel_performance-pP.png}
        \caption{Relative improvement with $N = 2^{10}$.}
    \end{subfigure}
    \caption{Comparison of four different implementations of the Fast Fr\'echet Distance algorithm on a laptop (CPU: i7-11800H, GPU: GeForce RTX 3080 Mobile (16\,GB VRAM), CUDA Version: 12.2) using 32-bit floating point numbers. \textit{Vanilla} and \textit{Linear} refer to Alg.~\ref{alg:no_recursion} and \ref{alg:linear}, respectively. The SIMD implementation uses a batch size of $B = 32$ (twice the size of a 512-bit register in order to improve the instruction level parallelism) and relies on the AVX-512 instruction set. The CUDA implementation uses a grid and block size of 128 and 64, respectively, that was measured to perform best for $N=2^{13}$ and $P=2^{10}$.}
    \label{fig:benchmark_laptop}
\end{figure}

The performance gain between Alg.~\ref{alg:vanilla} and \ref{alg:linear} is small but slightly increases when $P$ get large.
In contrast, the SIMD variant offers a constant improvement of roughly $19\times$ during the first and even up to $22\times$ during the second experiment.
We found that instruction level parallelism and an improved memory locality by packing $B=2 \times 16$ floating point numbers into 512-bit vector registers and using AVX-512 instructions explain this impressive boost.
Taking into account the large number of available threads of the GPU, the results of the CUDA implementation is less impressive and only outperforms the SIMD algorithm running on a single CPU core for large values of $N$.

@@

\section{Summary}
\label{sec:summary}
In this paper, we have introduced the Fast Fr\'echet Distance algorithm in Alg.~\ref{alg:fast}, a recursion-free algorithm that calculates the discrete Fr\'echet distance with a linear memory overhead and can be implemented in four lines of code.
Our algorithm can easily be adopted, e.g., to estimate the distances between batches of polygonal curves and a reference curve in parallel.
We have shown benchmarks of implementations that efficiently utilize SIMD vector registers on a CPU and the parallelization potentials of a GPGPU.
Besides a possible application to uncertain curves, we further conjecture that even for sophisticated approaches that, for instance, use heuristics to avoid computing the full (discrete) Fr\'echet distance when not needed, will still benefit from our results as they are probably still bottlenecked by those cases, where the heuristic fails \citep{buchin23,baldus17,buchin17b,duetsch17}.

>>>>

We benchmark the effectiveness of such a parallelization scheme by implementing Alg.~\ref{alg:fast} in \texttt{C++} using vectorization via SIMD instructions (CPU) and a CUDA implementation (GPGPU), and compare the performance with implementations of Alg.~\ref{alg:no_recursion} and \ref{alg:linear}.
Note that we intentionally not pick Alg.~\ref{alg:vanilla} for reference as this variant crashes due to its recursion nature quickly for larger trajectory sizes.

@@

All implementations use 32-bit floating point numbers to represent (intermediate) steps during calculation and run on an Ubuntu 22.04 machine using GCC-10 for the CUDA variant and GCC-11 for the rest.
The SIMD variant uses the \texttt{C++} extensions for parallelism as implemented in GCC-11~\citep{N4808}.
Builds where optimized using the flags \texttt{-O3}, \texttt{-DNDEBUG}, \texttt{-march=native}, and \texttt{-ffast-math}.
The reported numbers where taken after a warm-up phase during which the algorithm under investigation was run with the same data to minimize unwanted affects such as library loading, etc., during the measurements.

@@

\begin{figure}[htbp]
    \centering
    \begin{subfigure}{.49\textwidth}
        \includegraphics[width=.8\textwidth]{img/laptop/abs_performance-N.png}
        \caption{Absolute runtime with $P = 2^{10}$.}
    \end{subfigure}
    \begin{subfigure}{.49\textwidth}
        \includegraphics[width=.8\textwidth]{img/laptop/abs_performance-pP.png}
        \caption{Absolute runtime with $N = 2^{10}$.}
    \end{subfigure}\\
    \begin{subfigure}{.49\textwidth}
        \includegraphics[width=.8\textwidth]{img/laptop/rel_performance-N.png}
        \caption{Relative improvement with $P = 2^{10}$.}
    \end{subfigure}
    \begin{subfigure}{.49\textwidth}
        \includegraphics[width=.8\textwidth]{img/laptop/rel_performance-pP.png}
        \caption{Relative improvement with $N = 2^{10}$.}
    \end{subfigure}
    \caption{Comparison of four different implementations of the Fast Fr\'echet Distance algorithm on a laptop (CPU: i7-11800H, GPU: GeForce RTX 3080 Mobile (16\,GB VRAM), CUDA Version: 12.2) using 32-bit floating point numbers. \textit{Vanilla} and \textit{Linear} refer to Alg.~\ref{alg:no_recursion} and \ref{alg:linear}, respectively. The SIMD implementation uses a batch size of $B = 32$ (twice the size of a 512-bit register in order to improve the instruction level parallelism) and relies on the AVX-512 instruction set; the baseline implementation uses the same technique to calculate $\sum_{ij} d_{ij}$. The CUDA implementation uses a grid and block size of 128 and 64, respectively, that was measured to perform best for $N=2^{13}$ and $P=2^{10}$.}
    \label{fig:benchmark_laptop}
\end{figure}

Besides the four variants, we further benchmark a baseline implementation that calculates the sum of all entries of the distance matrix $d \in \mathbb{R}^{P \times Q}$ using the very same vectorization techniques as the SIMD variant.
This gives us a decent approximation of an upper limit for the expected performance for the implementations running on the CPU, i.e., the difference between this baseline and the SIMD variant is the overhead induced by evaluating Eq.~\eqref{eq:dynprogprob} instead of simply summing the elements $d_{ij}$.

From Figs.~\ref{fig:benchmark_laptop} we see that the performance gain between Alg.~\ref{alg:vanilla} and \ref{alg:linear} is small but slightly increases when $P$ gets large.
In contrast, the SIMD variant comes close to our baseline and offers a constant improvement of roughly $19\times$ during the first and even up to $22\times$ during the second experiment.
We found that instruction level parallelism and an improved memory locality by packing $B=2 \times 16$ floating point numbers into 512-bit vector registers and using AVX-512 instructions explain this impressive boost.
Taking into account the large number of available threads of the GPU, the results of the CUDA implementation is less impressive and only outperforms the SIMD algorithm running on a single CPU core for large values of $N$.

@@

\section{Summary}
\label{sec:summary}
In this paper, we have introduced the Fast Fr\'echet Distance algorithm in Alg.~\ref{alg:fast}, a recursion-free algorithm that calculates the discrete Fr\'echet distance with a linear memory overhead and can be implemented in four lines of code.
Our algorithm can easily be adopted, e.g., to estimate the distances between batches of polygonal curves and a reference curve in parallel.
We have shown benchmarks of implementations that efficiently utilize SIMD vector registers on a CPU and the parallelization potentials of a GPGPU.
Besides a possible application to uncertain curves, we further conjecture that even for sophisticated approaches that, for instance, use heuristics to avoid computing the full (discrete) Fr\'echet distance when not needed, will still benefit from our results as they are probably still bottlenecked by those cases where the heuristic fails \citep{buchin23,baldus17,buchin17b,duetsch17}.
